%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Beamer Presentation
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{beamer}
\mode<presentation> {
  \usetheme{default}
  \usecolortheme{orchid}
  \setbeamertemplate{footline}[frame number]
  \setbeamertemplate{navigation symbols}{}
}

\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{bm}
\usepackage{array}
\usetikzlibrary{shapes.geometric, arrows, positioning}

\newcommand{\argmin}{\mathop{\mathrm{argmin}}\limits}
\newcommand{\argmax}{\mathop{\mathrm{argmax}}\limits}
\newcommand{\dens}{\mathrm{dens}}
\newcommand{\A}{A}
\newcommand{\F}{F}

\title[]{Densest Community Search}
\subtitle{Query-anchored dense subgraphs in directed citation networks}

\author{The-Anh Vu-Le}
\institute[UIUC]{
University of Illinois Urbana-Champaign \\
\medskip
\textit{vltanh@illinois.edu}
}
\date{}

% Insert a divider frame at the start of each of the four top-level sections
% (Motivation and Problem Statements, Approaches, Experiments, Conclusion) so
% the audience sees the structural break clearly in the deck.
\AtBeginSection[]{
  \begin{frame}
    \centering
    \usebeamerfont{title}\insertsectionhead\par
  \end{frame}
}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Outline}
  \tableofcontents
\end{frame}

\section{Motivation and Problem Statements}

\begin{frame}{Motivation}
  \begin{itemize}
    \item Citation networks: nodes are papers, directed edges are citations
    \item Query: a paper $q$
    \item Goal: retrieve a ``good'' citation community around $q$
  \end{itemize}

  \vfill

  \textbf{Question} Can we search for a neighborhood that is dense, query-centered, and structurally constrained?
\end{frame}

\begin{frame}{Oracle access}
  Edge convention: $u \to v \;\Longleftrightarrow\; u$ cites $v$.

  \vfill

  The graph is available only through 1-hop neighborhood queries. Each query for a paper $v$ returns its outgoing references and its incoming citations.

  \vfill

  \textbf{Incoming-citation cap}
  \begin{itemize}
    \item incoming-edge queries can be more expensive than outgoing:
          \begin{itemize}
            \item reference list is usually small, cited-by list is
                  unlimited
            \item references read off the paper itself; cited-by needs a
                  maintained database
          \end{itemize}
    \item the cap limits how many incoming edges are fetched per
          queried node
  \end{itemize}
\end{frame}

\begin{frame}{Combinatorial problem}
  Given
  \begin{itemize}
    \item directed graph $G=(V,E)$
    \item query node $q \in V$
    \item minimum size $k \ge 3$
  \end{itemize}

  \vfill

  Solve
  \begin{align*}
    \max_{S \subseteq V} \quad &
    \frac{|E(S)|}{|S|(|S|-1)}              \\
    \mathrm{s.t.} \quad        &
    q \in S                                \\
                               & |S| \ge k
  \end{align*}

  \vfill

  where $E(S)=\{(u,v)\in E \mid u,v\in S\}$.
\end{frame}

\begin{frame}{Optional connectivity}
  \textbf{Rooted $\kappa$-edge-connectivity}

  \vfill

  For every selected node $v \in S \setminus \{q\}$:
  \[
    \text{there are at least } \kappa
    \text{ edge-disjoint paths from } q \text{ to } v
  \]
  in the underlying undirected graph.

  \vfill

  \textbf{Note}
  \begin{itemize}
    \item $\kappa = 0$ disables the constraint
    \item This will be enforced lazily in the solver by min-cut cuts.
  \end{itemize}
\end{frame}

\begin{frame}{Nonlinear fractional 0-1 program}
  Let
  \begin{itemize}
    \item $x_v \in \{0,1\}$ indicate $v \in S$
    \item $y_{uv} \in \{0,1\}$ indicate directed edge $(u,v)$ is internal
  \end{itemize}

  \begin{align*}
    \max \quad          &
    \frac{\sum_{(u,v)\in E} y_{uv}}
    {\left(\sum_{u\in V}x_u\right)
      \left(\sum_{u\in V}x_u - 1\right)}
    \\
    \mathrm{s.t.} \quad &
    x_q = 1                                                  \\
                        & \sum_{u\in V} x_u \ge k            \\
                        & y_{uv} \le x_u,\quad y_{uv}\le x_v
    \quad \forall (u,v)\in E
  \end{align*}

  \vfill

  \textbf{Note} The lower bound $y_{uv}\ge x_u+x_v-1$ is not needed because $y$ has positive objective coefficient.
\end{frame}

\section{Approaches}

\begin{frame}{Dinkelbach transform}
  For a density estimate $\lambda$, define
  \[
    Z(\lambda)
    =
    \max_S
    \left\{
    |E(S)| - \lambda |S|(|S|-1)
    \right\}.
  \]

  \vfill

  \textbf{Key property}
  \[
    \lambda^\star \text{ is optimal}
    \quad \Longleftrightarrow \quad
    Z(\lambda^\star)=0.
  \]

  \vfill

  \begin{itemize}
    \item If $Z(\lambda)>0$, an improving denser set exists
    \item Update $\lambda \gets |E(S)|/[|S|(|S|-1)]$
    \item Stop when the parametric subproblem cannot improve density
  \end{itemize}
\end{frame}

\begin{frame}{Dinkelbach loop}
  \begin{enumerate}
    \item \textbf{Initialize:} build a feasible set by BFS from $q$, prune to improve density, set $\lambda_1$ to its density, $t = 1$
    \item \textbf{Optimize:} solve the parametric integer subproblem at $\lambda_t$
    \item \textbf{Convergence check:} if $Z(\lambda_t) \le 0$, terminate and return $\lambda_t$
    \item \textbf{Update:} otherwise set $\lambda_{t+1}$ to the density of the new subgraph, $t \gets t+1$, return to step 2
  \end{enumerate}

  \vfill

  \textbf{Outer loop} handles the fractional objective.

  \vfill

  \textbf{Inner loop} solves the parametric problem by branch-and-price.
\end{frame}

\begin{frame}{Dinkelbach parametric subproblem}
  At parameter $\lambda$:
  \begin{align*}
    Z(\lambda) = \max \quad &
    \sum_{(u,v)\in E} y_{uv} - \lambda \left[ \left(\sum_{u\in V} x_u\right)^{\!2} - \sum_{u\in V} x_u \right] \\
    \mathrm{s.t.} \quad     &
    x_q = 1                                                                                                    \\
                            & \sum_{u\in V} x_u \ge k                                                          \\
                            & y_{uv} \le x_u,\quad y_{uv} \le x_v \quad \forall (u,v) \in E                    \\
                            & x \in \{0,1\}^{|V|},\ y \in \{0,1\}^{|E|}
  \end{align*}
\end{frame}

\begin{frame}{Linearizing the penalty}
  The Dinkelbach penalty is
  \[
    |S|(|S|-1)
    =
    \left(\sum_v x_v\right)^2 - \sum_v x_v.
  \]

  Since $x_v^2=x_v$ for binary variables:
  \[
    |S|(|S|-1)
    =
    2\sum_{u<v}x_u x_v.
  \]

  \vfill

  Introduce pair variables:
  \[
    w_{uv} \approx x_u x_v,
    \qquad
    w_{uv} \ge x_u+x_v-1.
  \]

  \vfill

  Because $w$ has negative coefficient, this lower bound is enough at integer solutions.
\end{frame}

\begin{frame}{Linearized parametric subproblem}
  At parameter $\lambda$:
  \begin{align*}
    Z(\lambda) = \max \quad &
    \sum_{(u,v)\in E} y_{uv} - 2\lambda \sum_{u<v} w_{uv}                                                \\
    \mathrm{s.t.} \quad     &
    x_q = 1                                                                                              \\
                            & \sum_{u\in V} x_u \ge k                                                    \\
                            & y_{uv} \le x_u,\quad y_{uv} \le x_v \quad \forall (u,v) \in E              \\
                            & w_{uv} \ge x_u + x_v - 1 \quad \forall u<v                                 \\
                            & x \in \{0,1\}^{|V|},\ y \in \{0,1\}^{|E|},\ w \in \{0,1\}^{\binom{|V|}{2}}
  \end{align*}
\end{frame}

\begin{frame}{Active set and frontier}
  \textbf{Maintained sets}
  \begin{itemize}
    \item $A$: active nodes in the restricted master problem
    \item $F$: frontier nodes discovered but not active
    \item subgraph induced by $A \cup F$ is kept up to date via oracle queries, so no internal edge is missed
  \end{itemize}

  \vfill

  \textbf{Initialization}
  \begin{itemize}
    \item BFS rooted at $q$: each dequeued node is oracle-queried, joins $A$, and its neighbors enter $F$
    \item loop terminates at $|A| = k$
  \end{itemize}
\end{frame}

\begin{frame}{Restricted Master Problem}
  For active set $A$:

  \begin{align*}
    \max \quad         &
    \sum_{(u,v)\in E(A)} y_{uv}
    -2\lambda\sum_{\substack{u,v\in A\\u<v}}w_{uv}         \\
    \mathrm{s.t.}\quad &
    x_q = 1                                                \\
                       & \sum_{u\in A}x_u \ge k
    \quad (\text{dual: }\pi)                               \\
                       & y_{uv}\le x_u,\quad y_{uv}\le x_v
    \quad \forall (u,v)\in E(A)                            \\
                       & w_{uv}\ge x_u+x_v-1
    \quad \forall u<v \in A                                \\
                       & 0\le x,y,w\le 1
  \end{align*}

  \vfill

  This LP gives the bound inside branch-and-bound.
\end{frame}

\begin{frame}{Pricing a frontier node}
  Let $\bar{x}$ be the RMP LP solution and $\pi$ the dual price of $\sum_{u \in A} x_u \ge k$.

  \vfill

  \textbf{Fractional degree of $f \in F$}
  \[
    d_{\mathrm{frac}}(f) = \sum_{v \in N(f) \cap A} \bar{x}_v
  \]

  \vfill

  \textbf{Reduced cost}
  \[
    \bar{c}_f
    = \underbrace{d_{\mathrm{frac}}(f)}_{\text{edge gain}}
    \;-\; \underbrace{2\lambda \sum_{v \in A} \bar{x}_v}_{\text{pair penalty}}
    \;-\; \underbrace{\pi}_{\text{size dual}}
  \]

  \vfill

  \footnotesize See appendix A1 (fractional degree) and A2 (reduced cost derivation).
\end{frame}

\begin{frame}{Column generation}
  \textbf{Pricing rule}
  \[
    f^\star = \argmax_{f\in F}\bar{c}_f.
  \]

  \vfill

  \begin{itemize}
    \item \underline{Case $\bar{c}_{f^\star} \le 0$} no frontier node improves the LP. The RMP objective is a valid upper bound.
    \item \underline{Case $\bar{c}_{f^\star} > 0$} promote frontier nodes into $A$ and resolve.
  \end{itemize}

  \vfill

  Add the top-$b$ frontier nodes by $\bar{c}_f$ per round, not one at a time:
  \[
    b =
    \mathrm{clamp}
    \left(
    \lfloor \rho |A| \rfloor,\ b_{\min},\ b_{\max}
    \right).
  \]
  $\rho \in (0, 1]$ scales the batch with $|A|$; $b_{\min}$ and $b_{\max}$ floor and cap it.

  \vfill

  \textbf{Interpretation} Column generation is also graph expansion.
\end{frame}

\begin{frame}{Branch-and-bound}
  Each B\&B node stores $V_1$ (fixed in) and $V_0$ (fixed out).

  \vfill

  At each node:
  \begin{enumerate}
    \item apply bounds, run column generation
    \item prune if LP bound cannot beat incumbent
    \item if $\bar{x}$ is integral, check as a candidate
    \item else branch on a fractional variable
  \end{enumerate}

  \vfill

  Root fixes $x_q = 1$.
\end{frame}

\begin{frame}{Branching rule}
  A fractional node is judged by its fractional internal degree.

  \vfill

  \[
    d_{\mathrm{frac}}(v)
    =
    \sum_{u\in N(v)\cap A}\bar{x}_u.
  \]

  \vfill

  \begin{itemize}
    \item If $d_{\mathrm{frac}}(v) < 2\lambda$:
          \begin{itemize}
            \item $v$ is a hanging node
            \item branch zero-first
          \end{itemize}
    \item Otherwise:
          \begin{itemize}
            \item $v$ is more core-like
            \item branch one-first
          \end{itemize}
  \end{itemize}

  \vfill

  \textbf{Reason} $2\lambda$ is the marginal pair penalty rate.

  \vfill

  \footnotesize See appendix A3.
\end{frame}

\begin{frame}{Triangle cuts}
  \textbf{Problem} LP relaxation can spread fractional $x$ mass across nodes while keeping pair variables $w_{ij}$ at zero (e.g., $x_u = x_v = 0.5$ and $w_{uv} = 0$).

  \vfill

  \textbf{Solution} Add constraints that forbid this pattern:
  \[
    x_u + x_v + x_w - w_{uv} - w_{vw} - w_{uw} \le 1.
  \]

  \vfill

  \textbf{Note}
  \begin{itemize}
    \item separated only on fractional nodes
    \item bounded number per round
    \item skipped if LP gap is small
  \end{itemize}
\end{frame}

\begin{frame}{$\kappa$-connectivity cuts}
  Lazy enforcement. On each integer candidate $S$, scan $t \in S \setminus \{q\}$ via $q \to t$ max-flow on the unit-capacity undirected support graph.

  \vfill

  At the first $t$ with flow $< \kappa$:
  \begin{itemize}
    \item $R$ = source-side of min-cut (residual-reachable from $q$)
    \item add cut inequality
          \[
            \sum_{(u,v) \in \delta(R)} y_{uv} \;\ge\; \kappa \, x_t
          \]
          where $\delta(R) = \{(u,v) \in E : \text{exactly one of } u, v \in R\}$
    \item re-solve
  \end{itemize}

  \vfill

  One cut per round; the next violation, if any, is caught on re-solve.

  \vfill

  \footnotesize See appendix A4.
\end{frame}

\begin{frame}{Pruning integer incumbents}
  Since $|S| \ge k$, integer solutions may exceed $k$. Greedily remove nodes whose removal strictly improves the parametric objective.

  \vfill

  \begin{itemize}
    \item candidate: node with the smallest internal degree (weakest edge contribution)
    \item never remove $q$
    \item stop at $|S| = k$ or when no single removal helps
    \item preserve connectivity when $\kappa > 0$
  \end{itemize}

  \vfill

  \textbf{Effect} stronger incumbent, smaller returned community.
\end{frame}

\begin{frame}{Practical considerations}
  \textbf{Early stopping}
  \begin{itemize}
    \item soft time limit: no-improvement cap, fires only when an incumbent exists
    \item hard time limit: absolute wall-clock cap, fires regardless
    \item node limit: cap branch-and-bound nodes per subproblem
    \item gap tolerance: terminate when LP-incumbent gap drops below threshold
  \end{itemize}

  \vfill

  With any limit active, the solver returns the best incumbent. The LP bound stays a valid upper envelope on the parametric optimum.
\end{frame}

\begin{frame}{Average-degree baseline}
  \textbf{Problem}
  \[
    \max_{\substack{S \ni q \\ S \subseteq V_{\mathrm{loc}}}} \frac{|E(S)|}{|S|}
  \]
  where $V_{\mathrm{loc}}$ is a BFS-bounded local subgraph around $q$.

  \vfill

  \textbf{vs BP}
  \begin{itemize}
    \item objective: avg degree, not pair density (the two are not monotone in each other)
    \item no size constraint $|S| \ge k$
    \item no $\kappa$-connectivity constraint
  \end{itemize}

  \vfill

  \footnotesize Algorithm: adapted Goldberg's $s$--$t$ max-flow algorithm (polynomial time)
  \begin{itemize}
    \item materialising the local subgraph up-front
    \item forcing $q \in S$ by keeping $q$ on the source side of the cut
  \end{itemize}
\end{frame}

\section{Experiments}

\begin{frame}{Exp 1: solver correctness on enumerable graphs}
  \textbf{Goal}
  \begin{itemize}
    \item check that the solver recovers the true optimum on graphs
          small enough to enumerate every subset
  \end{itemize}

  \vfill

  \textbf{Fixture}
  \begin{itemize}
    \item Erd\H{o}s--R\'enyi random graphs $G(n, p)$
    \item $n = 20$
    \item $p \in \{0.25,\, 0.50,\, 0.75\}$
    \item seed $\in \{ 0, 1, 2, 3, 4 \}$ (so 15 graphs total)
    \item query $q$: highest-degree node per graph
  \end{itemize}
\end{frame}

\begin{frame}{Exp 1: protocol}
  \textbf{Brute-force oracle}
  \begin{itemize}
    \item enumerate every $S \ni q$ with $|S| \ge k$
    \item for each $(k, \kappa) \in \{3,4,5\} \times \{0,1,2\}$, record the maximum-density $S$ with $\lambda(S) \ge \kappa$
    \item mark cells with no feasible $S$ as infeasible
  \end{itemize}

  \vfill

  \textbf{Solver}
  \begin{itemize}
    \item run on every (method, $k$, $\kappa$, seed) cell
    \item no time, node, gap, or Dinkelbach limit
    \item skip cells the brute force flagged infeasible (solver otherwise exhausts the search proving infeasibility; wall time blows up on sparse seeds)
  \end{itemize}
\end{frame}

\begin{frame}{Exp 1: result}
  \textbf{AvgDeg}
  \begin{itemize}
    \item exact match on $15/15$ graphs at every $p$
  \end{itemize}

  \vfill

  \textbf{BP}
  \begin{itemize}
    \item every returned $S$ satisfies $|S| \ge k$ and rooted $\kappa$-edge-connectivity ($30/30$ cells)
    \item edge density matches BF optimum on $30/30$ cells
    \item zero solver errors
  \end{itemize}
\end{frame}

\begin{frame}{Citation dataset and split}
  \textbf{Dataset} PyTorch Geometric's CitationFull Cora-ML
  \begin{itemize}
    \item 2,995 nodes
    \item 8,416 directed edges
    \item 7 paper labels
  \end{itemize}

  \vfill

  \textbf{Train pool} every other node ($2{,}191$ on Cora\_ML): provides labels for the majority-vote prediction rule. Never used as a query.

  \vfill

  \textbf{Query pool} pure sources with out-degree $\ge 2$:
  \begin{itemize}
    \item val: 402 queries
    \item test: 402 queries
  \end{itemize}

  \textbf{Hard subset} 59 test nodes where the depth-1 BFS vote is wrong.
\end{frame}

\begin{frame}{Baselines and BP}
  \begin{itemize}
    \item \textbf{BFS} outgoing-citation neighborhood at depth $d \in \{1, 2, 3\}$ ($d = 1$ is the cheapest baseline)
    \item \textbf{AvgDeg} exact densest local subgraph by average degree
    \item \textbf{BP} branch-and-price for directed edge density, swept over $k$ and $\kappa$
  \end{itemize}
\end{frame}

\begin{frame}{Exp 2: scope}
  \begin{itemize}
    \item intrinsic perspective: label-agnostic geometry of the
          returned community
    \item evaluated on val + test combined, so every queried node
          contributes
  \end{itemize}

  \vfill

  Grid:
  \begin{itemize}
    \item BFS depth $\in \{1, 2, 3\}$
    \item AvgDeg single cell
    \item BP $k \in \{3, 4, 5\}$, $\kappa \in \{0, 1, 2\}$
    \item one Gurobi seed for BP $(42)$; deterministic methods once
    \item soft no-improvement cap $60$ s, hard wall cap $300$ s on BP cells
  \end{itemize}
\end{frame}

\begin{frame}{Exp 2: cluster quality on val + test}
  Mean $\pm$ std over val + test (804 queries), Cora-ML, single Gurobi seed.

  \vfill

  \tiny
  \begin{center}
  \begin{tabular}{lcccccc}
    \toprule
    Method & size & $\lambda_2$ & ncut & cond. & avg.deg & dens. \\
    \midrule
    BFS $d=1$                & 4.9 $\pm$ 2.8   & 0.96 $\pm$ 0.30 & 1.23 $\pm$ 0.15 & 0.78 $\pm$ 0.18 & 1.17 $\pm$ 0.50 & 0.35 $\pm$ 0.11 \\
    BFS $d=2$                & 13.1 $\pm$ 11.5 & 0.40 $\pm$ 0.29 & 1.07 $\pm$ 0.11 & 0.77 $\pm$ 0.17 & 1.68 $\pm$ 0.69 & 0.21 $\pm$ 0.11 \\
    BFS $d=3$                & 24.6 $\pm$ 26.2 & 0.26 $\pm$ 0.29 & 1.05 $\pm$ 0.12 & 0.76 $\pm$ 0.17 & 1.86 $\pm$ 0.76 & 0.15 $\pm$ 0.11 \\
    AvgDeg                   & 21.2 $\pm$ 15.2 & 0.32 $\pm$ 0.38 & 0.70 $\pm$ 0.53 & 0.69 $\pm$ 0.16 & 3.02 $\pm$ 1.55 & 0.21 $\pm$ 0.12 \\
    BP $k{=}3$ $\kappa{=}0$  & 3.3 $\pm$ 0.7   & 1.23 $\pm$ 0.34 & 1.36 $\pm$ 0.26 & 0.81 $\pm$ 0.17 & 1.20 $\pm$ 0.42 & 0.51 $\pm$ 0.10 \\
    BP $k{=}3$ $\kappa{=}1$  & 3.3 $\pm$ 0.7   & 1.25 $\pm$ 0.28 & 1.39 $\pm$ 0.13 & 0.80 $\pm$ 0.17 & 1.19 $\pm$ 0.41 & 0.51 $\pm$ 0.10 \\
    BP $k{=}3$ $\kappa{=}2$  & 3.3 $\pm$ 0.7   & 1.27 $\pm$ 0.27 & 1.41 $\pm$ 0.11 & 0.81 $\pm$ 0.17 & 1.10 $\pm$ 0.44 & 0.48 $\pm$ 0.12 \\
    BP $k{=}4$ $\kappa{=}0$  & 4.4 $\pm$ 0.8   & 0.75 $\pm$ 0.48 & 0.94 $\pm$ 0.53 & 0.78 $\pm$ 0.16 & 1.59 $\pm$ 0.48 & 0.47 $\pm$ 0.09 \\
    BP $k{=}4$ $\kappa{=}1$  & 4.2 $\pm$ 0.8   & 0.92 $\pm$ 0.28 & 1.21 $\pm$ 0.10 & 0.80 $\pm$ 0.17 & 1.47 $\pm$ 0.42 & 0.46 $\pm$ 0.09 \\
    BP $k{=}4$ $\kappa{=}2$  & 4.2 $\pm$ 0.7   & 0.94 $\pm$ 0.28 & 1.24 $\pm$ 0.08 & 0.80 $\pm$ 0.17 & 1.34 $\pm$ 0.46 & 0.42 $\pm$ 0.12 \\
    BP $k{=}5$ $\kappa{=}0$  & 5.0 $\pm$ 0.7   & 0.51 $\pm$ 0.44 & 0.74 $\pm$ 0.57 & 0.77 $\pm$ 0.15 & 1.76 $\pm$ 0.50 & 0.44 $\pm$ 0.09 \\
    BP $k{=}5$ $\kappa{=}1$  & 5.1 $\pm$ 1.0   & 0.73 $\pm$ 0.27 & 1.14 $\pm$ 0.08 & 0.79 $\pm$ 0.17 & 1.67 $\pm$ 0.47 & 0.41 $\pm$ 0.09 \\
    BP $k{=}5$ $\kappa{=}2$  & 5.2 $\pm$ 1.1   & 0.72 $\pm$ 0.28 & 1.16 $\pm$ 0.08 & 0.79 $\pm$ 0.17 & 1.52 $\pm$ 0.54 & 0.37 $\pm$ 0.11 \\
    \bottomrule
  \end{tabular}
  \end{center}
\end{frame}

\begin{frame}{Exp 2: cluster quality reading}
  \begin{itemize}
    \item BP at small $k$: highest density and $\lambda_2$ (density $0.51$, $\lambda_2 \approx 1.25$).
    \item AvgDeg: best external isolation (ncut $0.70$, cond.\ $0.69$) and highest avg.deg by construction.
    \item BFS: dominated on internal density at every depth.
    \item Raising $k$ in BP trades density for size; $k{=}5$ $\kappa{=}0$ lands close to AvgDeg's ncut at much higher density.
  \end{itemize}

  \vfill

  \textbf{Claim} BP wins density and algebraic connectivity; AvgDeg wins external isolation.
\end{frame}

\begin{frame}{Exp 2: cost on val + test}
  Mean $\pm$ std over val + test (804 queries), Cora-ML, single seed. Distributions are long-tailed (timeouts on hardest queries).

  \vfill

  \tiny
  \begin{center}
  \begin{tabular}{lcccccc}
    \toprule
    Method & wall (s) & queries & BB nodes & LP solves & cuts & $t_{\text{LP}}$ (s) \\
    \midrule
    BFS $d=1$                & 0.015 $\pm$ 0.004   & 1.0 $\pm$ 0.0     & --                 & --                  & --                  & --              \\
    BFS $d=2$                & 0.016 $\pm$ 0.005   & 4.9 $\pm$ 2.8     & --                 & --                  & --                  & --              \\
    BFS $d=3$                & 0.019 $\pm$ 0.006   & 13.1 $\pm$ 11.5   & --                 & --                  & --                  & --              \\
    AvgDeg                   & 0.042 $\pm$ 0.034   & 202 $\pm$ 210     & --                 & --                  & --                  & --              \\
    BP $k{=}3$ $\kappa{=}0$  & 1.39 $\pm$ 13.24    & 179 $\pm$ 201     & 119 $\pm$ 807      & 129 $\pm$ 831       & 87 $\pm$ 311        & 0.83 $\pm$ 9.88 \\
    BP $k{=}3$ $\kappa{=}1$  & 1.47 $\pm$ 13.71    & 179 $\pm$ 201     & 121 $\pm$ 772      & 132 $\pm$ 796       & 90 $\pm$ 308        & 0.87 $\pm$ 10.4 \\
    BP $k{=}3$ $\kappa{=}2$  & 10.3 $\pm$ 53.1     & 180 $\pm$ 202     & 81 $\pm$ 429       & 1106 $\pm$ 5877     & 1107 $\pm$ 5868     & 6.26 $\pm$ 36.3 \\
    BP $k{=}4$ $\kappa{=}0$  & 9.2 $\pm$ 36.5      & 200 $\pm$ 210     & 867 $\pm$ 2239     & 915 $\pm$ 2292      & 491 $\pm$ 670       & 8.55 $\pm$ 34.8 \\
    BP $k{=}4$ $\kappa{=}1$  & 28.2 $\pm$ 79.7     & 200 $\pm$ 210     & 1549 $\pm$ 3593    & 1618 $\pm$ 3697     & 756 $\pm$ 1348      & 26.9 $\pm$ 76.8 \\
    BP $k{=}4$ $\kappa{=}2$  & 47.9 $\pm$ 103.3    & 200 $\pm$ 210     & 1363 $\pm$ 3061    & 3795 $\pm$ 9691     & 3300 $\pm$ 9483     & 44.1 $\pm$ 95.6 \\
    BP $k{=}5$ $\kappa{=}0$  & 66.3 $\pm$ 109.7    & 202 $\pm$ 210     & 3235 $\pm$ 4689    & 3345 $\pm$ 4807     & 1343 $\pm$ 1588     & 63.0 $\pm$ 105.3 \\
    BP $k{=}5$ $\kappa{=}1$  & 92.4 $\pm$ 130.6    & 202 $\pm$ 210     & 3799 $\pm$ 5311    & 3961 $\pm$ 5485     & 1992 $\pm$ 2825     & 88.4 $\pm$ 126.3 \\
    BP $k{=}5$ $\kappa{=}2$  & 96.0 $\pm$ 131.8    & 202 $\pm$ 210     & 3197 $\pm$ 4493    & 4409 $\pm$ 7318     & 3326 $\pm$ 6744     & 90.8 $\pm$ 126.7 \\
    \bottomrule
  \end{tabular}
  \end{center}

  \vfill

  \scriptsize BB = B\&B nodes; LP = LP solves; cuts = cuts added; $t_{\text{LP}}$ = LP-solve time.
\end{frame}

\begin{frame}{Exp 3: scope and prediction rule}
  \begin{itemize}
    \item extrinsic perspective
    \item tune on val, evaluate on test and on the BFS-hard test
          subset
  \end{itemize}

  \vfill

  For each val or test query $q$:
  \begin{itemize}
    \item retrieve a community around $q$ from the chosen method
    \item predict $q$'s label by a uniform majority vote over the training-labelled members of that community
    \item ties broken by ascending label so the result is
          reproducible
    \item fallback (rare):
          \begin{itemize}
            \item walk outward by BFS from $q$, forbidding val and test nodes as stepping stones, until a training neighbour is reached
            \item if no training label is reachable, use the global training-majority label
          \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Exp 3: tuning on val}
  Macro precision, recall, F1 on val (Cora-ML, 402 queries), single Gurobi seed.

  \vfill

  \scriptsize
  \begin{center}
  \begin{tabular}{llrrr}
    \toprule
    Method & params                       & P     & R     & F1    \\
    \midrule
    BFS    & $d = 1$                      & \underline{\textbf{0.836}} & \underline{\textbf{0.832}} & \underline{\textbf{0.826}} \\
    BFS    & $d = 2$                      & 0.802 & 0.800 & 0.796 \\
    BFS    & $d = 3$                      & 0.781 & 0.764 & 0.763 \\
    \midrule
    AvgDeg & ---                          & \underline{0.714} & \underline{0.540} & \underline{0.558} \\
    \midrule
    BP     & $k = 3,\ \kappa = 0$         & 0.797 & 0.787 & 0.782 \\
    BP     & $k = 3,\ \kappa = 1$         & 0.799 & 0.790 & 0.785 \\
    BP     & $k = 3,\ \kappa = 2$         & 0.801 & 0.794 & 0.785 \\
    BP     & $k = 4,\ \kappa = 0$         & 0.762 & 0.683 & 0.694 \\
    BP     & $k = 4,\ \kappa = 1$         & 0.808 & 0.789 & 0.789 \\
    BP     & $k = 4,\ \kappa = 2$         & \underline{0.827} & \underline{0.810} & \underline{0.811} \\
    BP     & $k = 5,\ \kappa = 0$         & 0.726 & 0.616 & 0.628 \\
    BP     & $k = 5,\ \kappa = 1$         & 0.772 & 0.744 & 0.741 \\
    BP     & $k = 5,\ \kappa = 2$         & 0.795 & 0.784 & 0.780 \\
    \bottomrule
  \end{tabular}
  \end{center}

  \vfill

  \scriptsize \underline{Underline} = per-family best per column; \textbf{bold} = overall best per column. BP soft cap $60$ s, hard cap $300$ s.
\end{frame}

\begin{frame}{Exp 3: evaluation on test}
  Best per-method val params applied to test split (Cora-ML, 402 queries), single Gurobi seed.

  \vfill

  \scriptsize
  \begin{center}
  \begin{tabular}{llrrr}
    \toprule
    Method & val-best params     & P     & R     & F1    \\
    \midrule
    BFS    & $d = 1$              & 0.876 & 0.852 & 0.856 \\
    AvgDeg & ---                  & 0.770 & 0.575 & 0.599 \\
    BP     & $k = 4,\ \kappa = 2$  & 0.818 & 0.815 & 0.811 \\
    \bottomrule
  \end{tabular}
  \end{center}

  \vfill

  \scriptsize BP val-best is $(k, \kappa) = (4, 2)$.
\end{frame}

\begin{frame}{Exp 3: hard subset motivation}
  The full test split mixes two regimes.

  \vfill

  \begin{itemize}
    \item easy regime: the immediate citations already carry the
          right label, so depth-1 BFS is hard to beat
    \item hard regime: the immediate citations carry the wrong label,
          so the retrieval has to look further
  \end{itemize}

  \vfill

  BFS-hard test subset:
  \begin{itemize}
    \item test nodes where the depth-1 BFS vote disagrees with the
          true label
    \item 59 nodes on Cora-ML
  \end{itemize}
\end{frame}

\begin{frame}{Exp 3: evaluation on the hard subset}
  Best per-method val params applied to BFS-hard test nodes (Cora-ML, 59 queries). Conditional diagnostic: subset selected by test labels, so BFS $d=1$ scores 0 by construction.

  \vfill

  \scriptsize
  \begin{center}
  \begin{tabular}{llrrr}
    \toprule
    Method & val-best params      & P     & R     & F1    \\
    \midrule
    BFS    & $d = 1$              & 0.000 & 0.000 & 0.000 \\
    AvgDeg & ---                  & 0.269 & 0.251 & 0.146 \\
    BP     & $k = 4,\ \kappa = 2$ & 0.174 & 0.282 & 0.190 \\
    \bottomrule
  \end{tabular}
  \end{center}

  \vfill

  \scriptsize BP val-best is $(k, \kappa) = (4, 2)$. Different BP cells score very differently on hard cases.
\end{frame}

\section{CiteSeer cross-check}

\begin{frame}{CiteSeer: pool}
  \begin{itemize}
    \item train: $3{,}319$; val: $455$; test: $456$
    \item pure sources with out-degree $\ge 2$, same protocol as Cora-ML
    \item BFS-hard subset: $15$ test nodes
  \end{itemize}
\end{frame}

\begin{frame}{CiteSeer: cluster quality on val + test}
  Mean $\pm$ std over val + test (911 queries), CiteSeer, single seed.

  \vfill

  \tiny
  \begin{center}
  \begin{tabular}{lcccccc}
    \toprule
    Method & size & $\lambda_2$ & ncut & cond. & avg.deg & dens. \\
    \midrule
    BFS $d=1$                & 3.7 $\pm$ 1.2  & 1.05 $\pm$ 0.27 & 1.30 $\pm$ 0.14 & 0.71 $\pm$ 0.21 & 0.91 $\pm$ 0.29 & 0.36 $\pm$ 0.10 \\
    BFS $d=2$                & 5.8 $\pm$ 3.1  & 0.66 $\pm$ 0.39 & 1.18 $\pm$ 0.14 & 0.71 $\pm$ 0.21 & 1.06 $\pm$ 0.35 & 0.27 $\pm$ 0.12 \\
    BFS $d=3$                & 6.8 $\pm$ 4.6  & 0.59 $\pm$ 0.43 & 1.17 $\pm$ 0.15 & 0.71 $\pm$ 0.21 & 1.07 $\pm$ 0.33 & 0.26 $\pm$ 0.13 \\
    AvgDeg                   & 5.2 $\pm$ 2.7  & 0.78 $\pm$ 0.45 & 1.22 $\pm$ 0.19 & 0.70 $\pm$ 0.22 & 1.12 $\pm$ 0.42 & 0.32 $\pm$ 0.13 \\
    BP $k{=}3$ $\kappa{=}0$  & 3.0 $\pm$ 0.3  & 1.24 $\pm$ 0.25 & 1.41 $\pm$ 0.09 & 0.73 $\pm$ 0.21 & 0.87 $\pm$ 0.22 & 0.43 $\pm$ 0.09 \\
    BP $k{=}3$ $\kappa{=}2$  & 3.1 $\pm$ 0.3  & 1.24 $\pm$ 0.25 & 1.41 $\pm$ 0.09 & 0.73 $\pm$ 0.21 & 0.86 $\pm$ 0.22 & 0.42 $\pm$ 0.09 \\
    BP $k{=}4$ $\kappa{=}0$  & 3.8 $\pm$ 0.5  & 0.93 $\pm$ 0.29 & 1.26 $\pm$ 0.11 & 0.71 $\pm$ 0.22 & 1.00 $\pm$ 0.26 & 0.36 $\pm$ 0.09 \\
    BP $k{=}4$ $\kappa{=}2$  & 3.8 $\pm$ 0.5  & 0.93 $\pm$ 0.30 & 1.27 $\pm$ 0.10 & 0.72 $\pm$ 0.22 & 0.98 $\pm$ 0.26 & 0.36 $\pm$ 0.09 \\
    BP $k{=}5$ $\kappa{=}0$  & 4.4 $\pm$ 0.9  & 0.78 $\pm$ 0.33 & 1.21 $\pm$ 0.14 & 0.71 $\pm$ 0.21 & 1.06 $\pm$ 0.31 & 0.33 $\pm$ 0.09 \\
    BP $k{=}5$ $\kappa{=}2$  & 4.4 $\pm$ 0.9  & 0.78 $\pm$ 0.34 & 1.22 $\pm$ 0.12 & 0.71 $\pm$ 0.21 & 1.04 $\pm$ 0.32 & 0.32 $\pm$ 0.10 \\
    \bottomrule
  \end{tabular}
  \end{center}

  \vfill

  \scriptsize $\kappa = 1$ rows match $\kappa = 0$ at every $k$ on CiteSeer (constraint not binding), omitted.
\end{frame}

\begin{frame}{CiteSeer: cost on val + test}
  Mean $\pm$ std over val + test (911 queries), CiteSeer, single seed.

  \vfill

  \tiny
  \begin{center}
  \begin{tabular}{lcccccc}
    \toprule
    Method & wall (s) & queries & BB nodes & LP solves & cuts & $t_{\text{LP}}$ (s) \\
    \midrule
    BFS $d=1$                & 0.02 $\pm$ 0.00  & 1 $\pm$ 0   & --              & --              & --              & --              \\
    BFS $d=2$                & 0.01 $\pm$ 0.00  & 4 $\pm$ 1   & --              & --              & --              & --              \\
    BFS $d=3$                & 0.02 $\pm$ 0.01  & 6 $\pm$ 3   & --              & --              & --              & --              \\
    AvgDeg                   & 0.01 $\pm$ 0.00  & 7 $\pm$ 5   & --              & --              & --              & --              \\
    BP $k{=}3$ $\kappa{=}0$  & 0.02 $\pm$ 0.01  & 7 $\pm$ 5   & 6 $\pm$ 12      & 8 $\pm$ 13      & 12 $\pm$ 18     & 0.00 $\pm$ 0.00 \\
    BP $k{=}3$ $\kappa{=}2$  & 0.35 $\pm$ 9.94  & 7 $\pm$ 5   & 7 $\pm$ 13      & 66 $\pm$ 1742   & 70 $\pm$ 1743   & 0.29 $\pm$ 8.85 \\
    BP $k{=}4$ $\kappa{=}0$  & 0.02 $\pm$ 0.01  & 7 $\pm$ 5   & 8 $\pm$ 17      & 11 $\pm$ 20     & 28 $\pm$ 44     & 0.00 $\pm$ 0.01 \\
    BP $k{=}4$ $\kappa{=}2$  & 1.33 $\pm$ 19.85 & 7 $\pm$ 5   & 9 $\pm$ 18      & 259 $\pm$ 3735  & 275 $\pm$ 3736  & 1.17 $\pm$ 17.6 \\
    BP $k{=}5$ $\kappa{=}0$  & 0.02 $\pm$ 0.06  & 7 $\pm$ 5   & 10 $\pm$ 25     & 13 $\pm$ 30     & 46 $\pm$ 80     & 0.01 $\pm$ 0.03 \\
    BP $k{=}5$ $\kappa{=}2$  & 1.02 $\pm$ 17.20 & 7 $\pm$ 5   & 12 $\pm$ 31     & 208 $\pm$ 3343  & 240 $\pm$ 3345  & 0.89 $\pm$ 15.3 \\
    \bottomrule
  \end{tabular}
  \end{center}

  \vfill

  \scriptsize CiteSeer is roughly $100\times$ cheaper than Cora-ML on every BP cell; hard caps almost never fire.
\end{frame}

\begin{frame}{CiteSeer: tuning on val}
  Macro precision, recall, F1 on val (CiteSeer, 455 queries), single Gurobi seed.

  \vfill

  \scriptsize
  \begin{center}
  \begin{tabular}{llrrr}
    \toprule
    Method & params               & P     & R     & F1    \\
    \midrule
    BFS    & $d = 1$              & 0.939 & 0.952 & 0.943 \\
    BFS    & $d = 2$              & \underline{0.945} & 0.947 & \underline{0.944} \\
    BFS    & $d = 3$              & 0.943 & 0.946 & 0.943 \\
    \midrule
    AvgDeg & ---                  & \underline{0.944} & \underline{0.948} & \underline{0.943} \\
    \midrule
    BP     & $k = 3,\ \kappa = 2$ & 0.939 & 0.952 & 0.942 \\
    BP     & $k = 4,\ \kappa = 2$ & \underline{\textbf{0.946}} & 0.947 & \underline{\textbf{0.945}} \\
    BP     & $k = 5,\ \kappa = 2$ & 0.943 & \textbf{0.949} & 0.944 \\
    \bottomrule
  \end{tabular}
  \end{center}

  \vfill

  \scriptsize $\kappa = 0,1$ rows omitted (within $\le 0.005$ of $\kappa = 2$ at every $k$). All three families within $0.005$ F1.
\end{frame}

\begin{frame}{CiteSeer: evaluation on test and hard subset}
  Best per-method val params applied to test (456 queries) and to BFS-hard test ($15$ queries).

  \vfill

  \scriptsize
  \begin{center}
  \begin{tabular}{llrrrrrr}
    \toprule
    & & \multicolumn{3}{c}{full test} & \multicolumn{3}{c}{BFS-hard ($n=15$)} \\
    \cmidrule(lr){3-5} \cmidrule(lr){6-8}
    Method & val-best params      & P     & R     & F1    & P     & R     & F1    \\
    \midrule
    BFS    & $d = 2$              & 0.938 & 0.948 & 0.942 & 0.000 & 0.000 & 0.000 \\
    AvgDeg & ---                  & 0.932 & 0.943 & 0.937 & 0.067 & 0.025 & 0.036 \\
    BP     & $k = 4,\ \kappa = 2$ & 0.938 & 0.946 & 0.942 & 0.000 & 0.000 & 0.000 \\
    \bottomrule
  \end{tabular}
  \end{center}

  \vfill

  \scriptsize Families tie within noise on full test. Hard subset is tiny ($n=15$); macro F1 is high-variance.
\end{frame}

\section{OpenAlex frontend}

\begin{frame}{From simulation to live citation graph}
  \begin{itemize}
    \item the experiments so far ran on a local CSV of Cora-ML
    \item the same solver speaks to OpenAlex, an open citation
          database
    \item one query returns a paper's outgoing references and (capped)
          incoming citations
    \item every node materialised by the solver is fetched live
    \item lazy oracle keeps API cost proportional to the papers BP
          pulls in
  \end{itemize}
\end{frame}

\begin{frame}{Web frontend}
  \begin{itemize}
    \item input: a paper identifier as the query
    \item active set and frontier evolve in real time as the solver runs
    \item dense community + boundary citations rendered as a graph
    \item per-paper metadata pulled live from OpenAlex
    \item fits as a reading-list builder around a single query paper
  \end{itemize}
\end{frame}

\section{Conclusion}

\begin{frame}{Main interpretation}
  \begin{itemize}
    \item \textbf{BP wins internal geometry.} Densest community at smallest size (density $\approx 0.5$, $\lambda_2 \ge 0.77$, ncut $\ge 1.20$). Tight cliques anchored at $q$.
    \item \textbf{AvgDeg wins external isolation and raw avg degree.} cond.\ $0.69$ and avg.deg $3.02$; pair density and $\lambda_2$ drop accordingly.
    \item \textbf{BFS wins classification on Cora-ML} because test papers are pure sources whose direct citations are in the train pool. Depth 1 F1 $= 0.856$ on full test.
    \item \textbf{On the BFS-hard slice the ranks flip.} BP $k = 4,\ \kappa = 2$ scores F1 $= 0.190$, AvgDeg $0.146$, BFS depth 1 $= 0$ by construction.
    \item \textbf{Cost is workload-heterogeneous.} BFS sub-ms; AvgDeg $\sim 40$ms; BP median fast but mean blows up at $\kappa = 2$ (timeouts on hardest queries).
  \end{itemize}

  \vfill

  \textbf{Claim} BP is a dense-community retriever. Classification F1 follows from the geometry, not the other way around.
\end{frame}

\begin{frame}{Limitations}
  \begin{itemize}
    \item Reported numbers cover Cora\_ML and CiteSeer. DBLP and Cora runs
          are in progress.
    \item The temporal split makes one-hop BFS unusually strong: test nodes
          are pure sources whose direct citations are in the train set.
    \item The hard subset is a conditional diagnostic on test queries where depth-1 BFS fails; BFS depth 1 scores 0 by construction.
    \item BP relies on Gurobi.
    \item Exact global optimality is over the discovered graph region; the
          oracle interface limits what BP can reason about.
    \item OpenAlex mode depends on API coverage and latency.
  \end{itemize}
\end{frame}

\begin{frame}{Next steps}
  \begin{itemize}
    \item Multi-dataset classification experiments
    \item Synthetic planted-community benchmarks
    \item Ablate:
          \begin{itemize}
            \item $\kappa$ cuts
            \item triangle cuts
            \item branching rule
            \item pruning
            \item batch pricing
          \end{itemize}
    \item Add latency-aware OpenAlex experiments
    \item Explore hybrid objectives:
          \[
            \text{density} + \text{label purity} + \text{semantic similarity}
          \]
  \end{itemize}
\end{frame}

\begin{frame}{Takeaways}
  \begin{itemize}
    \item Query-anchored dense community search is a different task from BFS classification. BP optimises internal density; BFS reads local label structure; AvgDeg picks the broadest densest set.
    \item BP returns the densest, tightest communities (density $0.5$, $\lambda_2 \ge 0.77$). AvgDeg dominates external isolation. BFS dominates classification F1 on Cora-ML by exploiting local-label structure.
    \item On the BFS-hard test slice BP narrowly beats AvgDeg ($0.190$ vs $0.146$); BFS depth 1 scores zero by construction.
    \item BP cost is long-tailed: median fast, tail dominated by hard-cap timeouts at $\kappa = 2$ and $k = 4$.
  \end{itemize}
\end{frame}

\begin{frame}{}
  \centering
  \Large
  \emph{Thank you.}
\end{frame}

\appendix

% ── A1. Fractional degree (refs: Pricing a frontier node) ────────────────────

\begin{frame}{A1. Fractional degree: directionality and cap}
  \textbf{Directionality}
  \begin{itemize}
    \item $N(v) = N^+(v) \cup N^-(v)$: undirected union of outgoing references and incoming citations
    \item reciprocal pairs $(u,v),(v,u) \in E$ contribute $v$ twice, matching directed edge density
  \end{itemize}

  \vfill

  \textbf{Robustness to the incoming-citation cap}
  \begin{itemize}
    \item every active node is oracle-queried, so its complete outgoing list is recorded
    \item ingestion is symmetric: writing $u \to v$ also marks $v \in \text{adj\_in}[u]$ and $u \in \text{adj\_in}[v]$
    \item every edge with at least one endpoint in $A$ is captured, regardless of the cap
  \end{itemize}

  \vfill

  Cap limits frontier coverage (which papers can ever join $A \cup F$), not the correctness of $d_{\mathrm{frac}}(v) \cap A$.
\end{frame}

% ── A2. Reduced cost derivation (refs: Pricing a frontier node) ──────────────

\begin{frame}{A2.1. Reduced cost: setup}
  RMP at parameter $\lambda$, active set $A$:
  \begin{align*}
    \max \quad        & \sum_{(u,v)\in E(A)} y_{uv} - 2\lambda \sum_{\substack{u<v\\u,v\in A}} w_{uv} \\
    \text{s.t.} \quad & x_q = 1                                                                       \\
                      & \sum_{u\in A} x_u \ge k \quad (\text{dual: } \pi)                             \\
                      & y_{uv} \le x_u,\; y_{uv} \le x_v                                              \\
                      & w_{uv} \ge x_u + x_v - 1                                                      \\
                      & x,y,w \in [0,1]
  \end{align*}

  LP optimum: $\bar{x}, \bar{y}, \bar{w}$. Size dual: $\pi$.

  \vfill

  \textbf{Question} If $f \in F$ joins $A$, does the LP improve?
\end{frame}

\begin{frame}{A2.2. Reduced cost: probe}
  Pretend $f$ joins $A$. New variables enter the LP:
  \begin{itemize}
    \item $x_f = 1$ (probe value)
    \item $y_{fv}$ for each edge $(f, v) \in E$ with $v \in A$
    \item $w_{fv}$ for each $v \in A$ (every pair, edge or not)
  \end{itemize}

  \vfill

  Freeze $\bar{x}_v$ for $v \in A$ at current LP values. Optimize new variables freely. The resulting objective change is the reduced cost:
  \[
    \bar{c}_f \;=\; \underbrace{(y\text{-gain})}_{\text{Step 1}} \;+\; \underbrace{(w\text{-penalty})}_{\text{Step 2}} \;+\; \underbrace{(\text{size correction})}_{\text{Step 3}}.
  \]
\end{frame}

\begin{frame}{A2.3. Reduced cost step 1: edge gain ($y$ terms)}
  For each new edge $(f, v) \in E$ with $v \in A$:

  \vfill

  \textbf{Bounds.}
  \[
    y_{fv} \le \min(x_f, x_v) = \min(1, \bar{x}_v) = \bar{x}_v.
  \]

  \textbf{Objective coefficient.} $+1$.

  \vfill

  Max problem, positive coefficient $\Rightarrow$ upper bound binds:
  \[
    y_{fv}^\star = \bar{x}_v.
  \]

  \vfill

  Total $y$-contribution:
  \[
    \sum_{v \in N(f) \cap A} \bar{x}_v \;=\; d_{\mathrm{frac}}(f).
  \]
\end{frame}

\begin{frame}{A2.4. Reduced cost step 2: pair penalty ($w$ terms)}
  For each $v \in A$ (every pair, edge or not):

  \vfill

  \textbf{Bound.}
  \[
    w_{fv} \ge x_f + x_v - 1 = \bar{x}_v.
  \]

  \textbf{Objective coefficient.} $-2\lambda$.

  \vfill

  Max problem, negative coefficient $\Rightarrow$ lower bound binds:
  \[
    w_{fv}^\star = \bar{x}_v.
  \]

  \vfill

  Total $w$-contribution:
  \[
    -2\lambda \sum_{v \in A} \bar{x}_v.
  \]
\end{frame}

\begin{frame}{A2.5. Reduced cost step 3: size dual ($\pi$ term)}
  Size constraint $\sum_{u \in A} x_u \ge k$ has dual price $\pi$.

  \vfill

  \textbf{Dual interpretation.} $\pi$ is the marginal LP value of relaxing $k$ by one unit.

  \vfill

  When $x_f = 1$, the LHS grows by one. The reduced cost picks up a shadow term $-\pi$.

  \vfill

  \begin{itemize}
    \item Size constraint slack at LP optimum: $\pi = 0$, no contribution.
    \item Size constraint binding ($|S| = k$): $\pi$ takes the sign that makes $-\pi$ the correct correction.
  \end{itemize}
\end{frame}

\begin{frame}{A2.6. Reduced cost: sum and decision}
  Sum the three contributions:
  \[
    \bar{c}_f \;=\; d_{\mathrm{frac}}(f) \;-\; 2\lambda \sum_{v \in A} \bar{x}_v \;-\; \pi.
  \]

  \vfill

  \textbf{Decision rule}
  \begin{itemize}
    \item $\bar{c}_f > 0$: adding $f$ strictly improves the LP $\Rightarrow$ promote to $A$.
    \item $\bar{c}_f \le 0$: no gain $\Rightarrow$ keep in $F$.
    \item $\max_{f \in F} \bar{c}_f \le 0$: no frontier node helps. Column generation terminates and the RMP LP objective is a valid upper bound on the parametric subproblem.
  \end{itemize}
\end{frame}

% ── A3. Branching threshold (refs: Branching rule) ───────────────────────────

\begin{frame}{A3. Branching threshold $2\lambda$}
  Dinkelbach objective:
  \[
    \sum_{(u,v)\in E} y_{uv} - 2\lambda \sum_{u<v} w_{uv}.
  \]
  Each pair carries penalty rate $2\lambda$.

  \vfill

  Marginal contribution of $v$ at fractional rate (per active neighbor $u$):
  \begin{itemize}
    \item edge gain: $+\bar{x}_u$ if $(u,v) \in E$
    \item pair penalty: $-2\lambda \bar{x}_u$ (every $u \in A$)
  \end{itemize}

  \vfill

  Aggregate: $v$ pays off when $d_{\mathrm{frac}}(v) \ge 2\lambda$.

  \vfill

  \begin{itemize}
    \item $d_{\mathrm{frac}}(v) \ge 2\lambda$: edge gain covers pair penalty rate $\Rightarrow$ core-like, branch one-first
    \item $d_{\mathrm{frac}}(v) < 2\lambda$: penalty dominates $\Rightarrow$ hanging, branch zero-first
  \end{itemize}

  \vfill

  Heuristic threshold; at LP optimum, included nodes typically satisfy $d_{\mathrm{frac}}(v) \ge 2\lambda$.
\end{frame}

% ── A4. $\kappa$-cut inequality (refs: $\kappa$-connectivity cuts) ───────────

\begin{frame}{A4. $\kappa$-cut inequality: derivation}
  \textbf{Menger's theorem (undirected)} For $t \in S$,
  \[
    \max(\text{edge-disjoint } q\text{-}t \text{ paths in } E(S)) \;=\; \min_{\substack{R \ni q \\ R \not\ni t}} |\delta(R) \cap E(S)|.
  \]
  So $S$ has $\ge \kappa$ edge-disjoint $q$-$t$ paths $\iff$ every $q$-$t$ cut $R$ has $\ge \kappa$ internal edges.

  \vfill

  \textbf{LP encoding} $y_{uv} = 1$ iff edge $(u,v)$ internal to $S$.
  \begin{itemize}
    \item $x_t = 1$: need $\sum_{(u,v) \in \delta(R)} y_{uv} \ge \kappa$
    \item $x_t = 0$: no requirement
  \end{itemize}
  Combine:
  \[
    \sum_{(u,v) \in \delta(R)} y_{uv} \;\ge\; \kappa \, x_t.
  \]

  \vfill

  \textbf{Validity} Holds at every $\kappa$-edge-connected integer $S$, so no feasible solution is excluded.
\end{frame}

\end{document}
